---
title: "Analyse the content of the "
output: html_document
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/Dropbox/threadAnalysis/working") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
library(ggplot2)
library(ggdendro)
library(ggraph)
# devtools::install_github("terminological/standard-print-output")
library(standardPrintOutput)
# devtools::install_github("terminological/tidy-info-stats")
devtools::load_all("~/Git/tidy-info-stats/")
# library(tidyinfostats)


rm(list = ls(all.names = TRUE))
devtools::load_all("~/Git/etl-analysis/r-packages/omop-utils/")
omop = Omop$new()
```

## Generate a test corpus from the first 50 orthopaedic and 50 rheumatology documents:

```{r}

searcher = omop$buildVocabSet()
searcher$applyUsedInFilter(omop$note,note_class_concept_id) %>% invisible()
searcher$toDataframe() %>% filter(concept_name %in% c("Rheumatology Note","Orthopaedic surgery note")) %>% select(concept_id, concept_name)
```

Use these concepts to pick the first 50 documents of each type (here we see thetop 10 codes in each group). This contains a lot of noise from single one off concepts

```{r}
rawDf = omop$note %>% filter(note_class_concept_id == 3032918L) %>% head(50) %>% union(
    omop$note %>% filter(note_class_concept_id == 3031179L) %>% head(50)
  ) %>%
  left_join(omop$note_nlp, by="note_id") %>%
  select(note_class_concept_id, note_id,lexical_variant, concept_id=note_nlp_concept_id) %>% 
  group_by(note_class_concept_id, note_id) %>% 
  #show_query()
  compute()

# summarise by note class
rawDf %>% 
  group_by(note_class_concept_id, concept_id) %>% 
  summarise(freq = n()) %>% ungroup() %>% 
  omop$getConceptNames() %>% 
  select(note_class_concept_name, concept_name, freq) %>% 
  group_by(note_class_concept_name) %>% arrange(desc(freq)) %>% filter(row_number() <= 10) %>% 
  collect() %>% ungroup() %>% group_by(note_class_concept_name) %>% 
  standardPrintOutput::saveTable("~/Dropbox/featureSelection/orthRheum/conceptsByCount")

localRawDf = rawDf %>% omop$getConceptNames() %>% collect()

```

And expand these concepts to include the hierarchy above them

```{r}
groupedDf = rawDf %>% omop$expandAncestorConcepts(concept_id) %>% compute()
localGroupedDf = groupedDf %>% omop$getConceptNames() %>%  collect()
# glimpse(groupedDf)

# localGroupedDf = groupedDf %>% collect()
groupedDf %>% group_by(note_class_concept_id, concept_id) %>% summarise(freq = sum(count)) %>%
  ungroup() %>% omop$getConceptNames() %>% select(note_class_concept_name, concept_name, freq) %>%
  group_by(note_class_concept_name) %>% arrange(desc(freq)) %>% filter(row_number() <= 10) %>% 
  ungroup() %>% group_by(note_class_concept_name) %>% 
  standardPrintOutput::saveTable("~/Dropbox/featureSelection/orthRheum/expandedConceptsByCount")
```

This table now contains a lot of potential noise from high level concepts.

## Calculate occurrence stats for concepts in the corpus binary MI stats both for plain cooccurence and for occurrence relative to okapi BM 25

TF-IDF scores to the rescue

* Calculate a corpus IDF based on all 100 documents
* Calculate a TFIDF score for each concept in each document

```{r}

# devtools::load_all("~/Git/tidy-info-stats/")
# groupedDf = groupedDf %>% group_by(note_class_concept_id, note_id)
idfDf = groupedDf %>% tidyinfostats::calculateIdf(concept_id, count)
localIdfDf = idfDf %>% omop$getConceptNames() %>% collect()

tfidfDf = groupedDf %>% tidyinfostats::calculateTfidf(concept_id, idfDf=idfDf ) %>% compute()

tfidfDf %>% 
  group_by(note_class_concept_id, concept_id) %>% 
  summarise(avg_bm25 = mean(norm_okapi_bm25)) %>% ungroup() %>% 
  omop$getConceptNames() %>% 
  select(note_class_concept_name, concept_name, avg_bm25) %>% 
  group_by(note_class_concept_name) %>% 
  arrange(desc(avg_bm25)) %>% 
  filter(row_number()<=10)  %>% collect() %>% ungroup() %>% 
  group_by(note_class_concept_name) %>% 
  standardPrintOutput::saveTable("~/Dropbox/featureSelection/orthRheum/expandedConceptsByOkapiBM25")

localTfidfDf = tfidfDf %>% omop$getConceptNames() %>%  collect()



#glimpse(tfidfDf)
```

calculate mutual information between concept_id and note_class_concept_id

This is using the binary presence or absence of the concept_id in the document to calculate the MI

N.B. N_t = f_y is probably correct here. N_t is the number of documents with that term in it.

<!-- ```{r} -->
<!-- #devtools::load_all("~/Git/tidy-info-stats/") -->
<!-- miDf = tfidfDf %>% ungroup() %>%  -->
<!-- 	tidyinfostats::probabilitiesFromCooccurrence(vars(note_class_concept_id), vars(concept_id)) %>%  -->
<!--   group_by(concept_id) %>%  -->
<!-- 	tidyinfostats::calculateMultiClassMI() %>% omop$getConceptNames() -->
<!-- glimpse(miDf %>% arrange(desc(I))) -->
<!-- ``` -->

```{r}
# devtools::load_all("~/Git/tidy-info-stats/")
miDf = groupedDf %>% ungroup() %>% group_by(concept_id) %>% calculateDiscreteBinaryMI(vars(note_class_concept_id), count)
localMiDf = miDf %>% omop$getConceptNames() %>%  collect()


miDf %>% omop$getConceptNames() %>% select(concept_name, I) %>% arrange(desc(I)) %>% filter(row_number()<=10) %>% ungroup() %>% 
  standardPrintOutput::saveTable("~/Dropbox/featureSelection/orthRheum/expandedConceptsByDiscreteMI")

```

and we can also calculate the MI between concept_id and note_class_concept_id using a continuous variable (in this case norm_okapi_bm25)



```{r}
devtools::load_all("~/Git/tidy-info-stats/")
# debug( calculateDiscreteContinuousPointwiseMI_KWindow)
miContinuousDf = tfidfDf %>% group_by(concept_id) %>% 
  tidyinfostats::calculateDiscreteContinuousMI(vars(note_class_concept_id), norm_okapi_bm25, method="KWindow")
# glimpse(miContinuousDf %>% arrange(I))
localMiContinuousDf = miContinuousDf %>% omop$getConceptNames() %>% collect()

miContinuousDf %>% omop$getConceptNames() %>% select(concept_name, I) %>% 
  arrange(desc(I)) %>% filter(row_number() <= 10) %>% ungroup() %>% 
  standardPrintOutput::saveTable("~/Dropbox/featureSelection/orthRheum/expandedConceptsByContinuousMI")
```

## Calculate top 30 terms co-occurrence

TODO: filter out those co-occurrences that are inevitable due to child parent relationships.

```{r}
#devtools::load_all("~/Git/tidy-info-stats/")
#debug(tidyinfostats::calculateCooccurrence)
options(error=NULL)
top50 = idfDf %>% inner_join(miContinuousDf %>% arrange(desc(I)) %>% head(50) %>% select(concept_id),by="concept_id")
cooc = groupedDf %>% tidyinfostats::calculateCooccurrenceMI(concept_id,top50) 

localCooc = cooc %>% omop$getConceptNames() %>%  collect()


tmp = cooc %>% filter(concept_id1 < concept_id2 ) %>% omop$getConceptNames() %>% select(concept_name1,concept_name2,I) 
tmp %>% arrange(desc(I))  %>% filter(row_number() <= 10) %>% 
  standardPrintOutput::saveTable("~/Dropbox/featureSelection/orthRheum/top10Coocurrences")
tmp %>% arrange(I)  %>% filter(row_number() <= 10) %>% 
  standardPrintOutput::saveTable("~/Dropbox/featureSelection/orthRheum/bottom10Coocurrences")
```

```{r}


graphCooc = groupedDf %>% omop$getSpanningGraphEdges(concept_id)
localGraphCooc = graphCooc %>% omop$getConceptNames() %>% collect()
```


G. D. Tourassi, E. D. Frederick, M. K. Markey, and C. E. Floyd Jr, “Application of the mutual information criterion for feature selection in computer-aided diagnosis,” Med. Phys., vol. 28, no. 12, pp. 2394–2402, Dec. 2001 [Online]. Available: http://dx.doi.org/10.1118/1.1418724

Has info about how to combine MIM networks, but it is quite old

TODO:

* Conditional MI from the co-occurrence network MI / Hx / Hy etc. and Conditional Mutual info between concepts okapiBM25 and outcome
* Look at the behaviour of this with respect to concept hierarchy position...


```{r}
# TODO this does not produce output for everything (needs at least 10 points per group)
# secondly method cannot give result for outcome when it does not appear - e.g. only positive associated scores are present - and nothing for negatives.
# this limits usefulness possibly as finding of absence in negative cases is a strong indication of usefulness, but a the moment this is not accounted for.
condMiContinuousDf = tfidfDf %>% group_by(concept_id) %>% 
  tidyinfostats::calculateDiscreteContinuousConditionalMI(vars(note_class_concept_id), norm_okapi_bm25, method="KWindow") %>% compute()

# TODO: this threw a mysterious error
localCondMiContinuousDf = condMiContinuousDf %>% omop$getConceptNames() %>% collect()

# selct all the edges with the unexpanded and expanded concept networks.

glimpse(condMiContinuousDf)
```



## Visualisation 

```{r}
# lets get this locally as too slow dealing with DB
# devtools::load_all("~/Git/etl-analysis/r-packages/omop-utils/")
# omop = Omop$new()

localRawEdges = rawDf %>% omop$getSpanningGraphEdges(concept_id) %>% omop$getConceptNames() %>% collect()
localGroupedEdges = groupedDf %>% omop$getSpanningGraphEdges(concept_id) %>% omop$getConceptNames() %>% collect()



```

```{r}

length(unique(localRawDf$concept_id))
length(unique(localRawEdges$ancestor_concept_id))
length(unique(localRawEdges$descendant_concept_id))

length(unique(localGroupedDf$concept_id))
length(unique(localGroupedEdges$ancestor_concept_id))
length(unique(localGroupedEdges$descendant_concept_id))

localIdfDf %>% group_by(f_t) %>% count()

localIdfDf %>% filter(f_t >= 9) %>% count()

```
```{r}
# mark nodes that were originally identified in the document as
# this is now a feature of Omop$expandConcepts so can be taken out


graphNodes = localIdfDf %>% ungroup() %>% arrange(concept_id) %>% mutate(id = row_number())
map = graphNodes %>% select(concept_id,id)

graphEdges = localGroupedEdges %>% 
  left_join(map %>% rename(descendant_concept_id=concept_id),by=c("descendant_concept_id")) %>% rename(from=id) %>%
  left_join(map %>% rename(ancestor_concept_id=concept_id),by=c("ancestor_concept_id")) %>% rename(to=id)
  

cooccurEdges = localCooc %>% filter(concept_id1 != concept_id2) %>% 
  left_join(
    graphEdges %>% rename(concept_id1 = ancestor_concept_id, concept_id2 = descendant_concept_id), by=c("concept_id1","concept_id2")
  ) %>% mutate(hierarchy = !is.na(ancestor_concept_name))

cooccurEdges %>% group_by(hierarchy) %>% count()

```

```{r}

graph = tidygraph::tbl_graph(nodes = graphNodes, edges = graphEdges, directed = TRUE)
# TODO: implement some kind of more generic filtering here
graph = graph %>% activate(nodes) %>% mutate(isRoot = node_is_sink(), score = f_t, isLabelled = FALSE)
graph = graph %>% activate(edges) %>% mutate(fromScore = .N()$f_t[from], toScore = .N()$f_t[to])

set.seed(101)
p = ggraph::ggraph(graph, layout="nicely")+
		ggraph::geom_edge_diagonal(aes(
						# colour=hsv(0,0,0.75*(1-toScore)),
						# alpha=stat(index)*(toScore-fromScore)+fromScore,
						# alpha=0.25+0.75*fromScore,
						# alpha=stat(index)*0.75*(fromScore-toScore)+0.75*toScore+0.25,
						edge_width=0.25+0.75*sqrt(fromScore)
				), alpha=0.1, flipped=TRUE)+
		ggraph::geom_node_point(aes(size=0.25+0.75*sqrt(score)),
						colour="#000000",
						alpha = 0.5,
						stroke=0)+
		ggraph::geom_node_point(aes(size=0.25+0.75*sqrt(score), filter=isRoot),
						colour="#ff0000",
						alpha = 1,
						stroke=0)+
		ggraph::geom_node_point(aes(size=0.25+0.75*sqrt(score), filter=isLabelled),
						colour="#0000ff",
						alpha = 1,
						stroke=0)+
		ggraph::geom_node_text(aes(
						label="", #ifelse(score>labelScore | isRoot,concept_name,NA),
						colour=ifelse(isRoot,"#ff0000","#0000ff")
				),
				alpha=0.8,
				repel=TRUE,
				size=(6/ggplot2:::.pt/(96/72)) #standardPrintOutput::labelInPoints(6)
		) + scale_fill_identity() + scale_color_identity() + scale_edge_color_identity() + scale_alpha_identity() +
		ggraph::theme_graph() + theme(legend.position = "none")

p
```


<!-- # a = Analyser$fromDataframe(omop,groupedDf) -->
<!-- # a$calculateTfidf() -->
<!-- # a$calculateOutcomeMI(note_class_concept_id) -->
<!-- cooccurMI = a$nodes %>% ungroup() %>% arrange(desc(mi)) %>% head(50) %>% a$calculateCooccurrenceMI() -->
<!-- cooccurCount = a$nodes %>% ungroup() %>% arrange(desc(f_t)) %>% head(50) %>% a$calculateCooccurrenceMI() -->

# Cluster and plot similarity matrix
<!-- plotClustering = function(df) { -->
<!-- 	#df = cooccurMI -->
<!-- 	tmp = df %>% select(concept_name1, concept_name2, mi) %>% collect() -->
<!-- 	simMat = stats::xtabs(mi ~ concept_name1 + concept_name2,tmp, na.action = na.pass) -->
<!-- 	simMat = ifelse(simMat==0,t(simMat),simMat) -->
<!-- 	distMat = 1-simMat/max(simMat,na.rm=TRUE) -->
<!-- 	clust = hclust(as.dist(t(distMat))) -->
<!-- 	dhc = as.dendrogram(clust) -->
<!-- 	ddata <- dendro_data(dhc, type = "rectangle") -->
<!-- 	p = ggplot(segment(ddata)) +  -->
<!-- 			geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +  -->
<!-- 			coord_flip() +  -->
<!-- 			scale_y_reverse() + -->
<!-- 			expand_limits(x = c(0, 1), y = c(-1.5, 0))+ -->
<!-- 			geom_text(data = label(ddata),  -->
<!-- 					aes(x = x, y = y, label = label),  -->
<!-- 					vjust = 0.5, hjust=0,nudge_y = 0.04, -->
<!-- 					#nudge_x = -0.1,#  -->
<!-- 					size=(6/ggplot2:::.pt/(96/72))) + -->
<!-- 			theme_dendro() -->
<!-- 	return(p) -->
<!-- 	# https://www.r-graph-gallery.com/339-circular-dendrogram-with-ggraph.html -->
<!-- 	# ggdendro::ggdendrogram(clust, rotate=TRUE, size=(6/ggplot2:::.pt/(96/72)))) -->
<!-- 	# https://www.data-to-viz.com/graph/edge_bundling.html -->
<!-- } -->

<!-- plotClustering(cooccurMI) -->
<!-- standardPrintOutput::saveHalfPageFigure(filename="~/Dropbox/threadAnalysis/vocab/MIDendrogram") -->

<!-- plotClustering(cooccurCount) -->
<!-- standardPrintOutput::saveHalfPageFigure(filename="~/Dropbox/threadAnalysis/vocab/CountDendrogram") -->

<!-- #### Table of concepts by max nPMI and MI ---- -->
<!-- nodes = a$nodes %>% omop$getConceptNames() %>% collect() -->
<!-- tmp = nodes %>% ungroup() %>% arrange(desc(mi)) %>% head(50) %>% select( -->
<!-- 		Class=note_class_concept_name, -->
<!-- 		Concept=concept_name, -->
<!-- 		MI = mi, -->
<!-- 		nPMI = npmi -->
<!-- 		) %>% group_by(Class)  -->
<!-- tmp %>% standardPrintOutput::saveTable(filename="~/Dropbox/threadAnalysis/vocab/TopTermsByMI") -->

<!-- ####  -->
<!-- # a$plotConceptNetwork(scoreVar=mi,labelScore=0.8) -->
<!-- # standardPrintOutput::saveFullPageFigureLandscape(filename="~/Dropbox/threadAnalysis/vocab/OrthRheumVocab") -->



<!-- p %>% standardPrintOutput::saveFullPageFigureLandscape(filename="~/Dropbox/threadAnalysis/vocab/OrthRheumVocab") -->