---
title: "Analyse the content of the "
output: html_document
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/Dropbox/threadAnalysis/working") })
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
devtools::load_all("~/Git/tidy-info-stats")
devtools::load_all("~/Git/standard-print-output")
# detach("package:standardPrintOutput", unload=TRUE)
library(standardPrintOutput)
theme_set(defaultFigureLayout())

# ---- database setup ----
setwd('~/Dropbox/threadAnalysis/summary')
config <- config::get(file="~/Dropbox/db.yaml")
omop <- Omop$new("omop",config)
```

```{r}
View(omop$note_nlp %>% select(lexical_variant,note_nlp_concept_id) %>% omop$getConceptNames() %>% head(1000) %>% collect())
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
