---
title: "Analyse the content of the "
output: html_document
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/Dropbox/threadAnalysis/working") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r include=FALSE}
library(tidyverse)
library(reshape2)
library(ggplot2)
# library(zoo)
library(extrafont)
# library(dbplot)
# detach("package:dbplot", unload=TRUE)
# devtools::install_github("terminological/standard-print-output")
# detach("package:standardPrintOutput", unload=TRUE)
library(standardPrintOutput)
theme_set(defaultFigureLayout())

# ---- database setup ----
setwd('~/Dropbox/threadAnalysis/summary')
config <- config::get(file="~/Dropbox/db.yaml")
omop <- Omop$new("omop",config)
```

## Visits

```{r}
visits <- omop$visit_occurrence %>% summarise(min=min(visit_start_date),max=max(visit_start_date)) %>% collect()
```


### People by year of birth (figure) and gender (table) ----

```{r}
totalPeople=(omop$person %>% count() %>% collect())$n[1]
tmp <- omop$person %>% group_by(gender_concept_id) %>%
  summarise(Count = n()) %>% collect() %>%
  omop$getConceptNames() %>% select(Gender=gender_concept_name, Count) %>%
  mutate(`% age`=Count/totalPeople*100) %>%
  union(tibble(Gender=factor("total"),Count=totalPeople,`% age`=100))
tmp %>% saveTable("peopleByGender",colWidths = c(0.5,0.25,0.25))
tmp
```

```{r}
tmp <- omop$person %>%
  filter(!is.na(year_of_birth) & gender_concept_id != 0 & year_of_birth>(2019-110)) %>%
  group_by(gender_concept_id,year_of_birth) %>%
  summarise(count=n()) %>% collect() %>% omop$getConceptNames()

p <- tmp %>%
  ggplot(aes(x=year_of_birth,y=count,fill=gender_concept_name))+
  xlab("Year of birth")+ylab("Count")+
  geom_col(width=0.7)+
  scale_fill_manual(name="Gender",values=c("MALE"="#b0b0ff","FEMALE"="#ffb0b0"),breaks=c("MALE","FEMALE"),labels=c("M","F"))+
  narrowAndTall()
p %>% saveHalfPageFigure("peopleByYob")
p
```

###  visits by type (table) and over time (figure)

```{r}
# omop$visit_occurrence %>%
#   inner_join(omop$concept, by=c("visit_concept_id"="concept_id")) %>%
#   dbplot_bar(concept_name) + scale_y_log10()
# saveThesisSixthPage("visitsByType")

# tryCatch({

tmp <- omop$person %>% left_join(
  omop$visit_occurrence %>%
    group_by(person_id, visit_concept_id) %>%
    summarise(count = n())) %>%
  collect() %>%
  omop$getConceptNames() %>%
  filter(!is.na(visit_concept_name)) %>%
  mutate(visit_concept_name=factor(visit_concept_name)) %>%
  complete(person_id,visit_concept_name,fill = list(count = 0)) %>%
  group_by("Visit type"=visit_concept_name) %>%
  summarise(
    "Count" = as.integer(sum(count)),
    "Per patient" = meanAndConfidence(count, f=threeDp)
  )
tmp %>% saveTable("visitsByType",colWidths = c(0.5,0.20,0.30))
tmp

tmp <- omop$visit_occurrence %>%
  group_by(visit_concept_id,visit_start_date) %>%
  summarise(count=n()) %>%
  select(visit_start_date,visit_concept_id,count) %>%
  collect() %>% omop$getConceptNames()
tmp <- tmp %>% mutate(roll = zoo::rollmedian(count,49, fill="extend"))
p <- tmp %>%
  ggplot(aes(x=visit_start_date, y=count,colour=visit_concept_name))+
  geom_point(alpha=0.2,size=1,stroke=0)+
  geom_line(aes(y=roll))+
  xlab("start date")+
  ylab("visits")+labs(colour=NULL)+scale_color_brewer(palette="Set1")+narrowAndTall()
p %>% saveThirdPageFigure("visitsByTime")
p
```

### non outpatient visits by time of week - los vs time of week (figure) ----

No clear cut weekly pattern but affected by deidentification & temporal resolution not good enough

```{r eval=FALSE}

# ---- Deprecated: non outpatient visits by type - length of stay (int days) and counts ----

# omop$visit %>%
#   mutate( los = datediff(minute,visit_start_datetime,visit_end_datetime)) %>%
#   filter( visit_concept_id != outpatient_visit_concept_id) %>%
#   group_by( visit_concept_id ) %>%
#   summarise(
#     count = n(),
#     mean = mean(los)
# ) %>% inner_join(omop$concept, by=c("visit_concept_id"="concept_id"))


tmp <- omop$visit_occurrence %>%
  mutate(
    los = datediff(minute,visit_start_datetime,visit_end_datetime),
    from_week_start = datediff(hour,dateadd(week, datediff(week, '19050101', visit_start_datetime), '19050101'),visit_start_datetime)
  ) %>%
  filter( visit_concept_id != local(concepts$outpatient_visit)) %>%
  group_by( from_week_start ) %>%
  summarise(
    count = n(),
    los = mean(los)
  ) %>%
  collect()

tmp <- tmp %>%
  mutate( los_trend = zoo::rollmean(los,21, fill="extend"))

p <- tmp %>%
  ggplot() +
  geom_line(aes(x=from_week_start, y=los_trend))+
  geom_point(aes(x=from_week_start, y=los))+
  xlab("hours from start of week")+
  ylab("daily average length of stay (minutes)")+
  scale_week_continuous()
p

# p %>% saveThirdPageFigure("losOverWeek")
```

### Maybe deprecated: non outpatient visits by time of week - los vs daily volume (figure)

No clear relationship between los and daily counts but again this will be affected by deidentification

https://www.statmethods.net/advstats/bootstrapping.html could be used if we wanted a confidence interval on Rsq

```{r}
tmp <- omop$visit_occurrence %>%
  mutate(
    los = datediff(minute,visit_start_datetime,visit_end_datetime)
  ) %>%
  filter( visit_concept_id != local(concepts$outpatient_visit)) %>%
  group_by( visit_start_date ) %>%
  summarise(
    count = n(),
    los = mean(los)
  )
p <- tmp %>% ggplot() +
  geom_point(aes(x=count, y=los),alpha=0.1)+
  xlab("daily volume of inpatient stays")+
  ylab("daily average length of stay (minutes)")
# p -> saveSixthPageFigure("losVsInptVisitsPerDay")
p

summary(lm(count ~ los,tmp))
```

### counts of various observations by day (figure)

```{r}
tmp <- omop$condition_occurrence %>% db_compute_count( condition_start_date ) %>%
  collect() %>% select( date = condition_start_date, count = `n()`) %>% mutate(type = 'condition') %>%
  union(
    omop$procedure_occurrence %>% db_compute_count( procedure_date ) %>%
      collect() %>% select( date = procedure_date, count = `n()`) %>% mutate(type = 'procedure')
  ) %>%
  union(
    omop$drug_exposure %>% db_compute_count( drug_exposure_start_date ) %>%
      collect() %>% select( date = drug_exposure_start_date, count = `n()`) %>% mutate(type = 'drug')
  ) %>%
  union(
    omop$note %>% db_compute_count( note_date ) %>%
      collect() %>% select( date = note_date, count = `n()`) %>% mutate(type = 'note')
  ) %>%
  union(
    omop$measurement %>% db_compute_count( measurement_date ) %>%
      collect() %>% select( date = measurement_date, count = `n()`) %>% mutate(type = 'measurement')
  ) %>%
  union(
    omop$observation %>% db_compute_count( observation_date ) %>%
      collect() %>% select( date = observation_date, count = `n()`) %>% mutate(type = 'observation')
  )
tmp <- tmp %>% arrange(date) %>% group_by(type) %>% mutate( count_trend = zoo::rollmedian(count,49, fill="extend"))

# https://www.ryanopel.com/post/dplyr-zoo-rollapply/
p <- ggplot(tmp %>% filter(date > '2000-01-01'), aes(x=date, y=count, colour=type))+
  annotate("rect", xmin = visits$min, xmax = visits$max, ymin = 5600, ymax = 5800)+
  geom_point(alpha=0.2,stroke=0,size=1)+
  geom_line(aes(y=count_trend))+
  lims(x=c(as.Date('2009-01-01'),as.Date('2019-02-01')),y=c(0,6000))+
  scale_color_brewer(palette="Set3")+narrowAndTall()

p %>% saveHalfPageFigure("dataPointsOverTime")
p
```

### summary of various observations by patient (table)

```{r}

counts <-
    omop$visit_occurrence %>% count() %>% mutate(table="Visits") %>%
    union(omop$condition_occurrence %>% count() %>% mutate(table="Condition codes")) %>%
    union(omop$procedure_occurrence %>% count() %>% mutate(table="Procedures")) %>%
    union(omop$drug_exposure %>% count() %>% mutate(table="Medication orders")) %>%
    union(omop$note %>% count() %>% mutate(table="Clinical note metadata")) %>%
    union(omop$note_nlp %>% count() %>% mutate(table="NLP derived codes")) %>%
    union(omop$measurement %>% count() %>% mutate(table="Measurements")) %>%
    union(omop$observation %>% count() %>% mutate(table="Observations")) %>% collect();

distincts <-
  omop$visit_occurrence %>% distinct(visit_concept_id) %>% count() %>% mutate(table="Visits") %>%
  union(omop$condition_occurrence %>% distinct(condition_concept_id) %>% count() %>% mutate(table="Condition codes")) %>%
  union(omop$procedure_occurrence %>% distinct(procedure_concept_id) %>% count() %>% mutate(table="Procedures")) %>%
  union(omop$drug_exposure %>% distinct(drug_concept_id) %>% count() %>% mutate(table="Medication orders")) %>%
  union(omop$note %>% distinct(note_class_concept_id) %>% count() %>% mutate(table="Clinical note metadata")) %>%
  union(omop$note_nlp %>% distinct(note_nlp_concept_id) %>% count() %>% mutate(table="NLP derived codes")) %>%
  union(omop$measurement %>% distinct(measurement_concept_id) %>% count() %>% mutate(table="Measurements")) %>%
  union(omop$observation %>% distinct(observation_concept_id) %>% count() %>% mutate(table="Observations")) %>% collect();

people = omop$person %>% count() %>% collect() %>% .$n
tmp <- counts %>% mutate(per_patient = twoDp(n/people)) %>% inner_join(distincts %>% rename(dist=n));
tmp %>% select("Data type"=table,"Total"=n,"Per patient"=per_patient, "Distinct types"=dist) %>%
  mergeCells() %>%
  huxtable::set_align(huxtable::everywhere,3,"right")
tmp %>% saveTable("dataPoints",colWidths = c(0.4,0.2,0.2,0.2))
tmp
```

```{r eval=FALSE}
# ---- los by visit start time density map - doesn't work as los is integer days
# omop$visit %>%
#   mutate(
#     log_los = log(datediff(minute,visit_start_datetime,visit_end_datetime)+1),
#     from_week_start = datediff(minute,dateadd(week, datediff(week, '19050101', visit_start_datetime), '19050101'),visit_start_datetime)
#   ) %>%
#   filter( visit_concept_id != 9202) %>%
#   db_compute_raster2(from_week_start,log_los) %>%
#   collect() %>%
#   ggplot() +
#   geom_rect(aes(xmin=from_week_start, xmax=from_week_start_2,ymin=log_los,ymax=log_los_2, fill=`n()`))
#
# omop$visit %>%
#   mutate(
#     los = datediff(minute,visit_start_datetime,visit_end_datetime),
#     from_week_start = datediff(minute,dateadd(week, datediff(week, '19050101', visit_start_datetime), '19050101'),visit_start_datetime)
#   ) %>%
#   filter( visit_concept_id != 9202) %>%
#   filter( los < 30*24*60) %>%
#   mutate( log_los = log(los+1)) %>%
#   db_compute_raster2(from_week_start,log_los) %>%
#   collect() %>%
#   ggplot() +
#   geom_rect(aes(xmin=from_week_start, xmax=from_week_start_2,ymin=log_los,ymax=log_los_2, fill=`n()`))
```

### notes by type summary (table)

```{r}

tmp <- omop$note %>%
  # group_by(note_type_concept_id,note_class_concept_id) %>% summarise(count = n()) %>%
  # inner_join(omop$concept_ancestor, by=c("note_type_concept_id"="descendant_concept_id")) %>%
  # inner_join(omop$concept, by=c("ancestor_concept_id"="concept_id")) %>%
  group_by(note_type_concept_id) %>%
  summarise(count = n()) %>%
  omop$getConceptNames() %>%
  collect() %>%
# tmp <- tmp %>% ungroup() %>%
# select(-note_type_concept_id,-note_class_concept_id) %>%
  select(`Note type`=note_type_concept_name,
         #`Note class`=note_class_concept_name,
         Count=count) %>%
  arrange(desc(Count))

tmp = tmp %>% bind_rows( tibble(`Note type`="Total",Count=sum(tmp$Count)) ) 
tmp %>% saveTable("noteTypeCounts")
tmp
```

### notes by class (figure)

```{r}

tmp <- omop$note %>% group_by(note_class_concept_id) %>% summarise(count = n()) %>%
  # inner_join(omop$concept_ancestor, by=c("note_class_concept_id"="descendant_concept_id")) %>%
  # inner_join(omop$concept, by=c("ancestor_concept_id"="concept_id")) %>%
  inner_join(omop$concept, by=c("note_class_concept_id"="concept_id")) %>%
  group_by(concept_name, note_class_concept_id) %>%
  summarise(count = sum(count)) %>%
  select(
    "Note class"=concept_name,
    # "Concept id"=note_class_concept_id,
    "Count" = count) %>%
  collect()

p <- tmp %>% ggplot(aes(area=`Count`,fill=`Note class`,label=`Note class`))+
  treemapify::geom_treemap(show.legend = FALSE, colour="black", start="topleft")+
  treemapify::geom_treemap_text(start="topleft", reflow=TRUE)+scale_fill_hue(h=c(180,270))
p %>% saveThirdPageFigure("noteClassCounts")
p
```

### notes by class over time (figure)

```{r}
tmp <- omop$note %>%
  inner_join(omop$concept, by=c("note_class_concept_id"="concept_id")) %>%
  mutate(concept_name=if_else(
    note_class_concept_id == 44803912 |
      note_class_concept_id == 36716202 |
      note_class_concept_id == 37395601,
    concept_name,
    "Clinical letter"
  )) %>% group_by(concept_name,note_date) %>%
  summarise(Count = n()) %>%
  select(Date = note_date, Type = concept_name, Count) %>%
  collect() %>%
  ungroup()

tmp <- tmp %>% arrange(Date) %>% group_by(Type) %>% mutate( count_trend = zoo::rollmedian(Count,49, fill="extend"))

# https://www.ryanopel.com/post/dplyr-zoo-rollapply/
ggplot(tmp, aes(x=Date, y=Count, colour=Type))+
  annotate("rect", xmin = visits$min, xmax = visits$max, ymin = 2800, ymax = 2900)+
  geom_point(alpha=0.2,stroke=0,size=1)+
  geom_line(aes(y=count_trend))+narrowAndTall()+
  lims(x=c(as.Date('2009-01-01'),as.Date('2019-02-01')),y=c(0,3000))+scale_color_brewer(palette="Set2")+narrowAndTall()

saveThesisHalfPage("noteClassesOverTime")
```

### TODO: diagnoses

```{r}

omop$note %>% group_by(note_title) %>% summarise(count = n()) %>% arrange(desc(count))
omop$note %>% group_by(note_source_value) %>% summarise(count = n()) %>% arrange(desc(count))
```

