---
title: "Test the functions of the omop search"
output: html_document
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/Dropbox/threadAnalysis/working") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# library(devtools)
# library(dplyr)
# library(tidygraph)
# library(ggraph)

omop$finalize()
devtools::load_all("~/Git/etl-analysis/r-packages/omop-utils/")
omop = Omop$new()
s = Searcher$new(omop)
```

```{r}
# conduct a simple search and return result in df
myocInf = Searcher$fromSearch(omop,"%myoc%inf%")
myocInf$summary()
```

# find used concepts in df

```{r}
measurementDf = Searcher$fromDataframe(omop$measurement, measurement_concept_id)
valuesDf = Searcher$fromDataframe(omop$measurement, value_as_concept_id)
```

# find concepts based on the code

```{r}
Searcher$fromConceptCode(omop,"%ICD%","J45%")
```

# find counts of different classes of concept

```{r}
measurementDf$getVocabularies()
measurementDf$getConceptClasses()
measurementDf$getDomains()
```

# find different classes of concept

```{r}
valuesDf$getRelationshipIds()
```

# investigate the co-occurence of tests and concept values

```{r}
labTestValues = omop$measurement %>% 
  filter(!is.na(value_as_concept_id)) %>% 
  group_by(measurement_concept_id,value_as_concept_id) %>% 
  summarise(count=n()) %>% 
  omop$getConceptNames() %>% 
  arrange(desc(count)) %>% collect()

testConcepts = Searcher$fromDataframe(labTestValues,measurement_concept_id)

valueConcepts = Searcher$fromDataframe(labTestValues,value_as_concept_id)

generalMicroMeasurement = testConcepts$clone()$applyFilter(
  (concept_name %like% "%culture%") |
  (concept_name %like% "%suscept%") |
  (concept_name %like% "%organis%") |
  (concept_name %like% "%immunoassay%"))
```

# understand lexical content of terms in used tests 

```{r}
generalRadiologyMeasurement = testConcepts$clone()$applyFilter(
  (concept_name %like% "CT%") |
  (concept_name %like% "MR%") |
  (concept_name %like% "US%") |
  (concept_name %like% "NM%") |
  (concept_name %like% "XR%") |
  (concept_name %like% "RF%") |
  (concept_name %like% "SPECT%"))
```

# load and save vocab sets

```{r}
generalMicroMeasurement$save("~/Dropbox/threadAnalysis/vocab/generalMicro")
generalMicroMeasurement = Searcher$load(omop,"~/Dropbox/threadAnalysis/vocab/generalMicro")

tmp = generalMicroMeasurement$clone()$expandAncestorConcepts(max=100)
tmp$summary()
tmp$applyConceptClassesFilter("LOINC Group")
```

# import the valueConcepts into a vocab set retaining the count data

```{r}
ts = omop$buildVocabSet()
ts$useConcepts(testConcepts %>% group_by(count))
ts$get() %>% filter(
  (concept_name %like% "%culture%") |
  # (concept_name %like% "%suscept%") |
  (concept_name %like% "%organis%")
) %>% ts$expandAncestorConcepts(min=1,max=1) %>% ts$show() %>% filter(
  (concept_name %like% "%culture%") |
  (concept_name %like% "%bacteria%")
) %>% ts$show()
```

```{r}
ts$search("Microbiology") %>% ts$applyVocabularyFilter("LOINC") %>% ts$expandDescendantConcepts(max=1) %>% ts$show()
ts$search("Antibiotic susceptibilities") %>% ts$expandDescendantConcepts() %>% ts$show()
```

View(labTestValues %>% filter(
  (measurement_concept_name %like% "%culture%") |
  (measurement_concept_name %like% "%organis%")
))

vs = omop$buildVocabSet()
# import the valueConcepts into a vocab set retaining the count data
vs$useConcepts(valueConcepts %>% group_by(count))
vs$show()
vs$get() %>% vs$applyFilters(c(
  concept_id != 4155143,
  concept_id != 4155142))




View(labTestValues %>% arrange(value_as_concept_id))


remoteLabTestValues = omop$con %>% copy_to(labTestValues)

# %>% collect()

omop$buildVocabSet()$findUsedConcepts(omop$measurement, value_as_concept_id)
omop$buildVocabSet()$findUsedConcepts(omop$measurement, measurement_concept_id)

micro = omop$buildVocabSet()$selectUsedConcepts(omop$measurement, value_as_concept_id)

omop$buildVocabSet()$findUsedConcepts(omop$observation, observation_concept_id)


s = omop$buildVocabSet()
s$selectConceptByCode("SNOMED","239913004")
s$expandAncestorConcepts(2)
s$get()

s$findRelationshipIds()

#### Anayser demo ----

omop$finalize()
devtools::document("~/Git/etl-analysis/r-packages/omop-utils/")
devtools::load_all("~/Git/etl-analysis/r-packages/omop-utils/")
omop = Omop$new()

groupedDf = omop$note %>% head(5) %>%
  left_join(omop$note_nlp, by="note_id") %>%
  select(note_id,lexical_variant, concept_id=note_nlp_concept_id) %>%
  omop$getConceptNames() %>% group_by(note_id) %>% compute()

a = Analyser$fromDataframe(omop,groupedDf)
groupedDf %>% a$countAncestorConcepts() %>% a$loadGraph()
a$plotConceptNetwork()

b = Analyser$new(omop)
groupedDf %>% b$countAncestorConcepts() %>% b$calculateTfidf() %>% a$loadGraph()
a$plotConceptNetwork(scoreVar = norm_okapi_bm25)


nodeIds = df %>% group_by(concept_id) %>% summarise(count = n())
compute(nodeIds)
nodes =
  nodeIds %>% inner_join(omop$concept, by="concept_id") %>% collect() %>%
  mutate(id = as.character(concept_id))
edges = omop$concept_ancestor %>% filter(min_levels_of_separation == 1) %>%
  inner_join(nodeIds, by=c("descendant_concept_id"="concept_id")) %>%
  inner_join(nodeIds, by=c("ancestor_concept_id"="concept_id")) %>%
  collect() %>%
  mutate(to=as.character(ancestor_concept_id),from=as.character(descendant_concept_id))
graph = tbl_graph(nodes = nodes, edges = edges, directed = TRUE)

plot(graph)

# c('star', 'circle', 'gem', 'dh', 'graphopt', 'grid', 'mds',
#  'randomly', 'fr', 'kk', 'drl', 'lgl')
root = graph %N>% filter(node_is_sink())
ggraph(graph, layout="lgl", root=root)+
geom_edge_fan(arrow = arrow(length=unit(0.1,"inches"))) +
# geom_conn_bundle(data = get_con(from = from, to = to,mode="out"), alpha = 0.1) +
geom_node_label(aes(label=concept_name))

graph %N>% filter(node_is_sink())


# create_notable('bull') %>%
#   activate(nodes) %>%
#   mutate(centrality = centrality_power()) %>%
#   activate(edges) %>%
#   mutate(mean_centrality = (.N()$centrality[from] + .N()$centrality[to])/2)


#' a = omop$analyser
#' df = s$findConceptByCode("SNOMED","239913004") %>% s$expandAncestorConcepts()
#' df %>% a$loadGraph()
