---
title: "Antibiotic resistance background"
output: html_document
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/Dropbox/antibioticResistance/output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
library(ggplot2)
library(tidyverse)
library(ensurer)
#devtools::install_github("terminological/standard-print-output")
#library(standardPrintOutput)
devtools::load_all("~/Git/standard-print-output/")

theme_set(standardPrintOutput::defaultFigureLayout())
# devtools::install_github("terminological/tidy-info-stats")
devtools::load_all("~/Git/tidy-info-stats/")
# library(tidyinfostats)


# rm(list = ls(all.names = TRUE))
devtools::load_all("~/Git/etl-analysis/r-packages/omop-utils/")
source("./antibioticResistanceSetup.R")
```

# Find bacterial species identified on cluture

* codes related to detection of bacteria from culture:
* codes defining the bacterial species that were identified 
* identify the sample type (UCUL etc.) - this is part of the measurement source value code

Plot species counts:

```{r}
speciesResults = omop$measurement %>% semi_join(bacteriaIdentified$result, by=c("measurement_concept_id"="concept_id")) %>% omop$getConceptNames() %>% collect() %>%
  mutate(sample = stringr::str_remove(measurement_source_value,"^.+\\|"))

# plot the summary of the species
speciesSummary = speciesResults %>% group_by(value_as_concept_name) %>% summariseTopN(n=8, sortVar=desc(count), count = n()) %>% mutate(order = row_number())
p=ggplot(speciesSummary, aes(x=reorder(value_as_concept_name,order),y=count))+geom_bar(stat="identity") + xlab("Organism")
p %>% saveSixthPageFigure("~/Dropbox/antibioticResistance/speciesFreq")

# TODO: plot species by sample type
# Need a grouped by summariseTopN function
```

Find test results for sensitivity or resistance.

These are tests whose results are concepts that are "Resistant" or "Sensistive" (but not "Methicillin resistant staph aureus")

* No matching results - relates to Flucloxacillin as Fluclox Susceptability missing from OMOP database as not licensed in US
* Sample from measurement source value
* Antibiotic name from concept string
* Summary counts of sensisitivity and resistance on per antibiotic basis and confidence intervals estimated from binomials


```{r}
#colnames(omop$measurement)
sensitivityResults = omop$measurement %>% semi_join(resistanceOutcome$result, by=c("value_as_concept_id"="concept_id")) %>% omop$getConceptNames() %>% collect() %>%
  mutate(
    sample = stringr::str_remove(measurement_source_value,"^.+\\|"),
    antibiotic=ifelse(measurement_concept_name=="No matching concept","Flucloxacillin",stringr::str_remove(measurement_concept_name,coll(" [Susceptibility]")))
  )
# glimpse(sensitivityResults)


# Generate summary stats and confidence interval
sensitivitySummary = sensitivityResults %>% 
  group_by(measurement_concept_name) %>% 
  summarizeTopN(
      n=12,
      sortVar=desc(total),
      total = n(),
      sensitive = sum(ifelse(value_as_concept_name=="Sensitive",1,0)),
      resistant = sum(ifelse(value_as_concept_name=="Sensitive",0,1))
  ) %>% mutate(
    #fracResist = resistant/total,
    mean = binom::binom.confint(resistant,total,methods="wilson")$mean,
    lower = binom::binom.confint(resistant,total,methods="wilson")$lower,
    upper = binom::binom.confint(resistant,total,methods="wilson")$upper
  ) %>% mutate(
    `Antibiotic` = stringr::str_remove(measurement_concept_name,coll(" [Susceptibility]")),
    `N` = total,
    `Resistance` = twoDp(mean*100, "%"),
    `95% CI (%)` = paste0(twoDp(lower*100),"\u2014",twoDp(upper*100))
  )

sensitivitySummary %>% ungroup() %>% select(`Antibiotic`,`N`,`Resistance`,`95% CI (%)`) %>% saveTable("~/Dropbox/antibioticResistance/resistFreq")
# or as a figure 
# ggplot(sensitivitySummary, aes(x = measurement_concept_name, y=mean, ymin=lower, ymax=upper)) + geom_point() + geom_errorbar()
```

# Visualise the monthly trend in susceptabilities

* Look at 

```{r}
sensByMonth = sensitivityResults %>% #filter(sample=="UCUL") %>% 
  mutate(ym = lubridate::floor_date(measurement_date,unit = "month")) %>% 
  group_by(ym,measurement_concept_name) %>% 
  summarize(
      total = n(),
      sensitive = sum(ifelse(value_as_concept_name=="Sensitive",1,0)),
      resistant = sum(ifelse(value_as_concept_name=="Sensitive",0,1))
  ) %>% mutate(
    #fracResist = resistant/total,
    mean = binom::binom.confint(resistant,total,methods="wilson")$mean,
    lower = binom::binom.confint(resistant,total,methods="wilson")$lower,
    upper = binom::binom.confint(resistant,total,methods="wilson")$upper
  ) %>% mutate(
    `Antibiotic` = stringr::str_remove(measurement_concept_name,coll(" [Susceptibility]")),
    `N` = total,
    `Resistance` = twoDp(mean*100, "%"),
    `95% CI (%)` = paste0(twoDp(lower*100),"\u2014",twoDp(upper*100))
  )

p3 = ggplot(sensByMonth %>% ungroup() %>% filter(
  `Antibiotic` %in% c("Amoxicillin","Amoxicillin+Clavulanate","Cefuroxime","Cephalexin","Ciprofloxacin","Erythromycin","Flucloxacillin","Gentamicin","Meropenem","Metronidazole","Nitrofurantoin","Penicillin","Teicoplanin","Tetracycline","Trimethoprim","Vancomycin")
), aes(x=ym,y=mean*100,ymin=lower*100,ymax=upper*100)) + geom_ribbon(alpha=0.2) + geom_point(size=0.5) + facet_wrap(vars(`Antibiotic`)) + xlab("time") + ylab("% resistant") 
p3 %>% saveHalfPageFigure("~/Dropbox/antibioticResistance/abxByTime")
```
# Resolve species that sensitivities relate to.

* there is no unifying id to match samples for species and sensitivities
* The combination of person_id, sample, and measuement_datetime is fairly unique unless there is multiple species grown from a single sample
* We exclude multiple species at this point as we cannot know which species sensitivities are for.


```{r}
# this pattern works for events occuring in the lead up to something.
# tmp = sensitivityResults %>% left_join(speciesResults, by=c("person_id","sample"), suffix=c(".sens",".org")) %>% filter(
#    measurement_datetime.org < measurement_datetime.sens & 
#    measurement_datetime.org > (measurement_datetime.sens - lubridate::days(10))
# )

# filter out any duplicates - make sure this is less  than 20% of total
speciesResultsUq = speciesResults %>% group_by(person_id,sample,measurement_datetime) %>% mutate(uniq = n()) %>% filter(uniq == 1) %>%
  ensure_that(nrow(.) > 0.8*nrow(speciesResults))

# get sensitivities by test result
sensByOrg = sensitivityResults %>% left_join(speciesResults, by=c("person_id","sample","measurement_datetime"), suffix=c(".sens",".org")) %>% 
  select(
    person_id,
    sensitivity=value_as_concept_name.sens,
    sample,
    antibiotic,
    organism=value_as_concept_name.org
  ) 

# generate 
matrix = sensByOrg %>% group_by(antibiotic,organism) %>% summarise(
  total = n(), 
  resistant = sum(ifelse(sensitivity=="Sensitive",0,1))
) %>% mutate(
  mean = binom::binom.confint(resistant,total,methods="wilson")$mean,
  lower = binom::binom.confint(resistant,total,methods="wilson")$lower,
  upper = binom::binom.confint(resistant,total,methods="wilson")$upper
)

top20orgs = matrix %>% group_by(organism) %>% summarise(total = sum(total)) %>% arrange(desc(total)) %>% head(20)
top20abx = matrix %>% group_by(antibiotic) %>% summarise(total = sum(total)) %>% arrange(desc(total)) %>% head(20)

matrix = matrix %>% filter(antibiotic %in% top20abx$antibiotic & organism %in% top20orgs$organism)
# TODO: N dimensional summaryTopN function woudl be useful here can be done with 2 group wise summaryTopN

p4 = ggplot(matrix, aes(x=reorder(antibiotic,-mean), y=reorder(organism,-mean), fill=mean*100))+geom_tile()+scale_fill_gradient2(low="green", mid="grey95",high="red",midpoint=5,na.value="black")+xlab(NULL)+ylab(NULL)+labs(fill="% resistant")+theme(
  panel.grid.major = element_blank(),panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text.x=element_text(angle = 60, hjust = 1)
)

p4 %>% saveHalfPageFigure("~/Dropbox/antibioticResistance/abxByOrg")

# TODO: these pictures would be better if we could have a hierarchy of drug types and organisms on both axes.
# need to use top20orgs, and lookup in omop concept list.
# grab a concept hierarchy?
# find the minimum spanning tree in concept hierarchy??
# see ?stats::heatmap

```

Also present this in terms of sensitivity by sample

```{r}
matrix2 = sensByOrg %>% group_by(antibiotic,sample) %>% summarise(
  total = n(), 
  resistant = sum(ifelse(sensitivity=="Sensitive",0,1))
) %>% mutate(
  mean = binom::binom.confint(resistant,total,methods="wilson")$mean,
  lower = binom::binom.confint(resistant,total,methods="wilson")$lower,
  upper = binom::binom.confint(resistant,total,methods="wilson")$upper
)

matrix2 = matrix2 %>% filter(antibiotic %in% top20abx$antibiotic)

p5 = ggplot(matrix2, aes(x=reorder(antibiotic,-mean), y=reorder(sample,total), fill=mean*100))+geom_tile()+scale_fill_gradient2(low="green", mid="grey95",high="red",midpoint=5,na.value="black")+xlab(NULL)+ylab(NULL)+labs(fill="% resistant")+theme(
  panel.grid.major = element_blank(),panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text.x=element_text(angle = 60, hjust = 1)
)

p5 %>% saveHalfPageFigure("~/Dropbox/antibioticResistance/abxBySample")

```

